\chapter*{Acknowledgments}

This work would not have been possible without the support of my family, colleagues, and friends. Firstly, I wish to thank my parents, family, and my partner Yarelis for their unwavering support. A special thanks to my advisor, Professor Sudhir Malik, who was fundamental to my academic growth, providing me with guidance and opportunities to shape me into what I am today. Thanks to Dr.~Scarlet Norberg (UPRM Post-Doc) for her continued support, guidance, and patience throughout this work. Scarlet also provided invaluable guidance even during my undergrad years at UPRM.

Thanks to the scientists and staff at the Fermilab LPC, in particular, Dr.~Gabriele Benelli for his phenomenal patience, leadership, and guidance throughout the years and during my visits at Fermilab. During the Emerging Jets analysis, I had a great professional opportunity to interact with scientists from the University of Maryland ( Dr. Sarah Eno, Dr. Long Wang, and Yi-Mu Chen), Fermilab (Dr. Kevin Pedro), and the University of Colorado Boulder (Dr. Kevin Stenson, Dr. Keith Ulmer, Dr. Alexx Perloff, Dr. Bill Ford, and Claire Savard) with special thanks to Dr.~Alexx Perloff for his guidance and mentoring.


I also wish to sincerely thank Fermilab LPC, the CMS Collaboration, CERN, and the Physics Department at the University of Puerto Rico at Mayagüez (UPRM) for providing me with facilities to conduct this research.
I would like to thank the National Science Foundation (NSF) for funding this work through the grant PHY-2111134 on "Physics Beyond Standard Model with the CMS Pixel Detector".

A lot of people have also contributed to my professional development which encompasses software training and broader impacts. A very special thanks to the NSF funded IRIS-HEP institute at Princeton University. While being a software training instructor at IRIS-HEP training programs I met several wonderful colleagues --  Kilian Lieret, Michel Villanueva, Wouter Deconick, Jim Pivarski, Henry Schreiner, Alexander Moreno -- who introduced me to state-of-the-art software in HEP.

I would also like to thank the Pathways to CU Physics Program and the University of Michigan - CERN REU program, respectively, for giving me a summer internship opportunity at CU Boulder and full semester scientific work at CERN.

This experience has provided me with several opportunities to disseminate my experience to high-school teachers and students in Puerto Rico and the international HEP community. Specifically, I would like to thank CROEM teacher Danelix Cordero who provided me initial opportunities for outreach work.

% A special thanks to all the wonderful people I have met in my early career as an aspiring physicist in HEP.
% Among these are those I met through the HEP Software Foundation (HSF) and IRIS-HEP whom have been influential for my professional and academic training in analysis software. In particular, Kilian Lieret, Michel Villanueva, Wouter Deconick, Jim Pivarski, Henry Schreiner, Alexander Moreno and many more that I cannot possibly fit into to these pages.
